
title: "Will it Rain Tomorrow in Australia?"
authors: Sofia Pope Trogu, Fairouz Baz Radwan, Auriane Mahfouz
date: "2023-07-15"
output:
  pdf_document: 
    toc: true
    toc_depth: 2
    number_sections: true
    keep_tex: yes
    fig_caption: yes
latex_engine: pdflatex
editor_options: 
  markdown: 
    wrap: sentence
---

# Import R Libraries

```{r import, eval = FALSE}
library(corrplot)
library(ggplot2)
library(caret)
library(magrittr)
library(gridExtra)  
library(scales)
library(DMwR2)
library(UBL)
library(caret)
library(MASS)
library(ipred)
library(rsample)
library(mlr)
library(knitr)
library(glmnet)
library(outliers)
library(class)
library(kableExtra)
library(tidyverse)
library(pROC)
library(dplyr)
```

# Data Preparation 

## Download the Rain Dataset

We downloaded the Rain Australia dataset as a CSV from the following link: <https://www.kaggle.com/datasets/jsphyg/weather-dataset-rattle-package/code?datasetId=6012&searchQuery=visual>.
The dataset is originally composed of 23 columns and 145,460 examples.
The aim of the data is to use available information about today's weather, i.e.
Temperature, Humidity, Pressure, to predict whether it will rain tomorrow.
Therefore, we originally start with 22 features and 1 target variable, RainTomorrow with binary classification (Yes=1, No=0).

```{r data loading}
file_path <- "/Users/Sofia/Desktop/Rain_Australia/weatherAUS.csv"
rain <- read.csv(file_path)
head(rain)
```

```{r, eval=FALSE}
summary(rain)
dim(rain)
```


## Variable Table of Contents
```{r load variable TOC}
# file_path2 <- "/Users/Sofia/Desktop/Rain_Australia/Feature_TOC2.csv"
# feature_toc <- read.csv(file_path2)
# 
# feature_toc %>%
#   knitr::kable(
#     format = "latex",
#     align = "l",
#     booktabs = TRUE,
#     longtable = TRUE,
#     linesep = "",
#     caption = "Table of features with description and units."
#     ) %>% row_spec(0,bold=TRUE) %>% 
#   kableExtra::kable_styling(
#       position = "left",
#       latex_options = c("striped", "repeat_header"),
#       stripe_color = "gray!15"
#     )

file_path3 <- "/Users/Sofia/Desktop/Rain_Australia/metrics_summary.csv"
metrics_summary <- read.csv(file_path3)

options(knitr.kable.NA = '')
metrics_summary %>%
  knitr::kable(
    format = "latex",
    align = "l",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    caption = "Table of performance metrics by model."
    ) %>% row_spec(0,bold=TRUE) %>% 
  kableExtra::kable_styling(
      position = "left",
      latex_options = c("striped", "repeat_header"),
      stripe_color = "gray!15"
    )
```